---
title: "Processing"
author: "Paige Duffin"
date: "11/3/2019"
output: html_document
---

---
title: "Pisaster Abundance"
output:
  html_document:
    df_print: paged
---

```{r message=FALSE}
library(caret) # Classification and Regression Training
library(corrplot)
library(dplyr) # A Grammar of Data Manipulations
library(forcats) # Tools for Working with Categorical Variables (Factors)
library(gdata) # Various R Programming Tools for Data Manipulation
library(ggplot2) # Create Elegant Data Visualizations Using the Grammar of Graphics
library(ggrepel) # Automatically Position Non-Overlapping Text Labels with ‘ggplot2’
library(ggridges) # Ridgeline Plots in ‘ggplot2’
library(ggthemes) # Extra Themes, Scales, and Geoms for ‘ggplot2’
library(gmodels) # Various R Programming Tools for Model Fitting
library(grid) # The Grid Graphics Package
library(gridExtra)
library(knitr) # A General-Purpose Package for Dynamic Report Generation in R
library(lmtest) # Testing Linear Regression Models
library(plyr) # Tools for Splitting, Applying and Combining Data
library(purrr) # Functional Programming Tools
library(readr) # Read Rectangular Text Data
library(rpart)
library(rattle)
library(scales) # Scale Functions for Visualization
library(shiny) # Web Application Framework for R
library(skimr) # Compact and Flexible Summaries of Data
library(stringr) # Simple, Consistent Wrappers for Common String Operations
library(tibble) # Simple Data Frames
library(tidyr) # Easily Tidy Data with ‘spread()’ and ‘gather()’ Functions
library(tidyverse) # Install and Load ‘Tidyverse’ Packages
library(usmap)
library(rworldmap)
library(ggmap)
library(maps)
library(mapdata)
library(cowplot)
library("googleway")
library("ggspatial") 
library("rnaturalearth") 
library("rnaturalearthdata")
theme_set(theme_bw())
library("sf")
```

```{r}
raw_SS_count <- read.csv("../../data/raw_data/seastarkat_size_count_totals_download.csv")
SS_count <- raw_SS_count
```

```{r}
SS_count$group_code <- as.character(SS_count$group_code)
SS_count$site_code <- as.character(SS_count$site_code)
SS_count$marine_site_name <- as.character(SS_count$marine_site_name)
SS_count$marine_season_code <- as.character(SS_count$marine_season_code)
SS_count$season_name <- as.character(SS_count$season_name)
SS_count$target_assemblage <- as.character(SS_count$target_assemblage)
SS_count$method_code <- as.character(SS_count$method_code)
SS_count$species_code <- as.character(SS_count$species_code)
SS_count$mpa_designation <- as.character(SS_count$mpa_designation)
SS_count$mpa_region <- as.character(SS_count$mpa_region)
SS_count$georegion <- as.character(SS_count$georegion)
SS_count$bioregion <- as.character(SS_count$bioregion)
SS_count$state_province <- as.character(SS_count$state_province)
SS_count$island <- as.character(SS_count$island)
SS_count$last_updated <- as.character(SS_count$last_updated)
glimpse(SS_count)
```

```{r}
skim(SS_count)
```

View all variables
```{r}
names(SS_count)
```

Examine each variable- characters first
```{r}
unique(SS_count$group_code)
SS_count %>% ggplot2::ggplot() + geom_bar(aes(group_code))
```
Hmmm, very biased towards UCSC. Maybe I should make an additional group_code that is UCSC vs Other
```{r}
SS_count1 <- SS_count %>% mutate(group_code_UCSC_other = recode(group_code, "CSUF" = "Other", "FEIRO" = "Other", "OCNMS" = "Other", "PBNERR" = "Other", "SSSC" = "Other", "UCLA" = "Other", "UCSB" = "Other", "UW" = "Other"))
skim(SS_count1$group_code_UCSC_other)
SS_count1$group_code <- as.factor(SS_count1$group_code)
SS_count1$group_code_UCSC_other <- as.factor(SS_count1$group_code_UCSC_other)
```

Now `site_code`
```{r}
unique(SS_count1$site_code)
SS_count1 %>% ggplot2::ggplot() + geom_bar(aes(site_code))
SS_count2 <- SS_count1
SS_count2$site_code <- as.factor(SS_count1$site_code)
```

Now `marine_site_name`
```{r}
SS_count3 <- SS_count2
unique(SS_count3$marine_site_name)
SS_count3 %>% ggplot2::ggplot() + geom_bar(aes(marine_site_name))
SS_count3$marine_site_name <- as.factor(SS_count3$marine_site_name)
```

One site name has a weird space after it- fixing that
```{r}
SS_count3 <- SS_count3 %>% mutate(marine_site_name = recode(marine_site_name, "McDonald Cove; Hood Canal " = "McDonald Cove; Hood Canal"))
#unique(SS_count3$marine_site_name) #double check to make sure that worked
SS_count3$marine_site_name <- as.factor(SS_count3$marine_site_name)
```

Now `marine_season_code`
```{r}
SS_count4 <- SS_count3
unique(SS_count4$marine_season_code)
SS_count4 %>% ggplot2::ggplot() + geom_bar(aes(marine_season_code))
SS_count4$marine_season_code <- as.factor(SS_count4$marine_season_code)
```
Kinda weird to have this var, when this info is already contained in the `season_name`, and `marine_common_year` vars, but it could be useful, and there appear to be no issues with it, so I'll leave it for now.....

Well, it does show information in this graph that is very interesting and allows for a timeline at a finer scale than by year. Actually, this looks very informative! 

Now `season_name`
```{r}
SS_count5 <- SS_count4
unique(SS_count5$season_name)
SS_count5 %>% ggplot2::ggplot() + geom_bar(aes(season_name))
SS_count5$season_name <- as.factor(SS_count5$season_name)
```

Now `target_assemblage`
```{r}
SS_count6 <- SS_count5
unique(SS_count6$target_assemblage)
SS_count6 %>% ggplot2::ggplot() + geom_bar(aes(target_assemblage))
```

Okay, so this analysis revealed that this variable is completelly uninformative, because its always the same target. Deleting this variable.  
```{r}
SS_count6 <- SS_count6[-c(12)]
#names(SS_count6) #just checking
```

Now `method_code`
```{r}
SS_count7 <- SS_count6
unique(SS_count7$method_code)
SS_count7 %>% ggplot2::ggplot() + geom_bar(aes(method_code))
```

Same story as `group_code`... very disproportionate. Doing the same thing as before- adding a new variable that is the most common method () vs "Other"
```{r}
SS_count7 <- SS_count7 %>% mutate(method_code_IP_other = recode(method_code, "TS30" = "Other", "GSES" = "Other", "BT25" = "Other"))
skim(SS_count7$method_code_IP_other)
SS_count7$method_code_IP_other <- as.factor(SS_count7$method_code_IP_other)
SS_count7$method_code <- as.factor(SS_count7$method_code)
```

Now `mpa_designation`
```{r}
SS_count8 <- SS_count7
unique(SS_count8$mpa_designation)
SS_count8 %>% ggplot2::ggplot() + geom_bar(aes(mpa_designation))
```

These designations aren't intuitive... let me see what the code says.. 
> `mpa_designation`: Describes whether the referenced site is located within a Marine Protected Area (MPA) or is a reference site. If this field is blank, the referenced site is not located within an MPA, or is not a reference site for an existing MPA site.

That doesn't really help me understand these levels because I don't know what a reference site is (even after a Google investigation). If this ends up being an important predictor in my model later, I will try to understand better what these categories mean. 

Changing to a factor and moving on.
```{r}
SS_count8$mpa_designation <- as.factor(SS_count8$mpa_designation)
```

Now `mpa_region`
```{r}
SS_count9 <- SS_count8
unique(SS_count9$mpa_region)
SS_count9 %>% ggplot2::ggplot() + geom_bar(aes(mpa_region))
```

Shouldn't the number of `NULL` entries in this variable (`mpa_region`) match the number of `NULL` entries in the last variable I looked at (`mpa_designation`)? What's going on? 
```{r}
NULL_mpa_designation <- SS_count9 %>% select(names(SS_count9)) %>% filter(mpa_designation == "NULL")
NULL_mpa_designation %>% ggplot2::ggplot() + geom_bar(aes(mpa_region))

NULL_mpa_region <- SS_count9 %>% select(names(SS_count9)) %>% filter(mpa_region == "NULL")
NULL_mpa_region %>% ggplot2::ggplot() + geom_bar(aes(mpa_designation))
```

I don't understand why some of the values for `mpa_designation` are labeled as not associated with an MPA (`NULL`) but their corresponding `mpa_region` is associated with a region. 

Because I see that all "NULL" `mpa_region`s are also "NULL" in the `mpa_designation` variable, but not all "NULL" `mpa_designations`s are also "NULL" in the `mpa_region` variable, it seems, at the very least, the `mpa_region` category is more informative. And, on top of that, these categories make sense to me, unlike those of the `mpa_designation` variable. Therefore, I'm making the decision to remove the `mpa_designation` variable from all downstream analyses. I have plenty of interesting variables to still work with (but if, for some reason, nothing is interesting in the models I run, I can always come back and add this variable back in, although it seems unlikely that `mpa_designation` would be an interesting predictor if I'd already determined that `mpa_region` weren't). 

Removing the `mpa_designation` variable and changing `mpa_region` to a factor. 
```{r}
# names(SS_count9) # determined that `mpa_designation` var is column 17
SS_count10 <- SS_count9[-c(17)]
SS_count10$mpa_region <- as.factor(SS_count10$mpa_region)
# names(SS_count10) #just checking that it worked
```

Now `georegion`
```{r}
SS_count11 <- SS_count10
unique(SS_count11$georegion)
SS_count11 %>% ggplot2::ggplot() + geom_bar(aes(georegion))
```

Very disproportionate.. but informative, and I don't know of a way to make it more evenly spread. 
```{r}
SS_count11$georegion <- as.factor(SS_count11$georegion)
```

Now `bioregion`
```{r}
SS_count12 <- SS_count11
unique(SS_count12$bioregion)
SS_count12 %>% ggplot2::ggplot() + geom_bar(aes(bioregion))
```

Some of these names are really long.. I'm going to abbreviate where I can. 
```{r}
SS_count12 <- SS_count12 %>% mutate(bioregion = recode(bioregion, "Government Point to Mexico" = "GovtPt_2_Mexico", "WA Olympic Coast to San Francisco" = "OlyCstWA_2_SanFran", "San Francisco to Government Point" = "SanFran_2_GovtPt", "Channel Islands South" = "ChannelIsl_South", "WA Salish Sea" = "SalishSea_WA", "Alaska to British Columbia" = "AK_2_BritColumb"))
# unique(SS_count12$bioregion) #double checking
SS_count12$bioregion <- as.factor(SS_count12$bioregion)
```

Now `state_province`
```{r}
SS_count13 <- SS_count12
unique(SS_count13$state_province)
SS_count13 %>% ggplot2::ggplot() + geom_bar(aes(state_province))
```

Just going to change these to the state abbreviations- personal preference. Also, there are no provinces in this variable, so renaming the variable to make more sense. 
```{r}
SS_count13 <- SS_count13 %>% mutate(state = recode(state_province, "Alaska" = "AK", "Washington" = "WA", "Oregon" = "OR", "California" = "CA"))
#unique(SS_count13$state) #double checking
SS_count13$state <- as.factor(SS_count13$state)

# Removing old variable
# names(SS_count13) # determined that `state_province` var is column 20
SS_count14 <- SS_count13[-c(20)]
# names(SS_count14) #just checking that it worked
```

Now `island`
```{r}
SS_count15 <- SS_count14
unique(SS_count15$island)
SS_count15 %>% ggplot2::ggplot() + geom_bar(aes(island))
```

What does "NULL" mean in this situation? 
```{r}
NULL_island <- SS_count15 %>% select(names(SS_count15)) %>% filter(island == "NULL")
NULL_island

#NULL_mpa_designation %>% ggplot2::ggplot() + geom_bar(aes(mpa_region))
```

Its always at the same site: "McDonald Cove; Hood Canal" (site_code: "HOO"), so it must mean SOMETHING.. but what? Let me look at the documentation. It says, "The name of the island where the referenced site is located. Sites not on islands are designated as mainland." 

Okay, so there's no good reason why this site has "NULL". Let me see if there are entries at this site where it isn't "NULL". 
```{r}
HOO_site <- SS_count15 %>% select(names(SS_count15)) %>% filter(site_code == "HOO")
unique(HOO_site$island)
```
Okay, so its only ever been "NULL" for this site in the `island` variable. I'll have to figure out for myself if this site is on an island or not.... googling now... 

Oh wow! Through this, I discovered the site that I downloaded this data from has detailed information on these collection sites... [here](https://marine.ucsc.edu/sitepages/mcdonaldcove.html). May use this information later... but for now, I used google maps to find this location. Its on the mainland- specifically on the Olympic Penninsula. 
Okay, cool. Mystery solved. Recoding it as "mainland." 
```{r}
SS_count15 <- SS_count15 %>% mutate(island = recode(island, "NULL" = "Mainland"))
unique(SS_count15$island)
SS_count15$island <- as.factor(SS_count15$island)
```

Now `last_updated`
```{r}
SS_count16 <- SS_count15
unique(SS_count16$last_updated)
SS_count16 %>% ggplot2::ggplot() + geom_bar(aes(last_updated))
```

Okay, as suspected, its completely uninformative as a future predictor variable. Removing it. 
```{r}
# names(SS_count16) # determined that `last_updated` var is column 21
SS_count17 <- SS_count16[-c(21)]
# names(SS_count17) #just checking that it worked
```

## Integers
First, `marine_sort_order`. I don't even know what this means, intuitively. What does the documentation say? 

"Used to order the sites geographically along the coast. Actual values are arbitrary, and should not be used for site identification."

Oh, interesting. That could actually be useful! 

```{r}
SS_count18 <- SS_count17
unique(SS_count18$marine_sort_order)
SS_count18 %>% ggplot2::ggplot() + geom_histogram(aes(marine_sort_order))
```

```{r}
unique_sites_code <- subset(SS_count18, !duplicated(SS_count18$marine_sort_order)) 
# skim(unique_sites_code)
# names(unique_sites_code) # want to keep columns 2, 3, 4, 5, 6
unique_sites_code1 <- unique_sites_code[c(2,3,4,5,6)]
#unique_sites_code1
ordered_sites_code_key <- unique_sites_code1
```

^^ Can use that data frame later to decode! 
